Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-03-11 Manuscript/HI_Model_ms.Rmd
updated manuscript files

 Carbon flow rates for each state type and moisture zone were estimated as the ratio of the IBIS-derived flux to the size of the originating carbon stock at each age [@sleeter_effects_2018]. A spatially explicit stationary growth multiplier was applied to each simulation cell to reflect local variations in net primary productivity (NPP) driven by microclimate. This spatial growth multiplier was the NPP anomaly for each cell relative to mean values for each combination of state type and moisture zone [@sleeter_effects_2019] calculated using empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008]. Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under each of two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017]. The impact of future changes in rainfall and temperature on NPP were represented by annual growth multipliers calculated using empirical NPP models [@schuur_productivity_2003; @del_grosso_global_2008] and climate model projections of temperature and rainfall for each radiative forcing scenario. The effect of future warming on turnover rates of dead organic matter were represented by temporal decay multipliers calculated using Q10 functions and climate model temperature projections for each radiative forcing scenario. We applied a Q10 of 2.0 for wood and soil organic matter decay flows [@kurz_cbm-cfs3:_2009; @sleeter_effects_2019] and a Q10 of 2.17 for litter decay flows [@bothwell_leaf_2014]. Transition-triggered carbon flows resulting from disturbances associated with land use change, timber harvesting, and wildfire were based on values from @don_impact_2011, @selmants_baseline_2017, and @daniel_integrating_2018.

@@

### *4.2 Comparison to other studies*
There are few estimates of contemporary ecosystem carbon balance for the main Hawaiian Islands, and even fewer model projections of future ecosystem carbon balance in response to climate and land use change.  Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2020 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019]. In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.314 Tg C y^-1^; @selmants_baseline_2017], despite our current model estimates of contemporary NPP and R~h~ being 9.5% and 15% lower, respectively.

@@

This study was funded by the U.S. Geological Survey Biological Carbon Sequestration Program. Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software, Nicholas Koch of Forest Solutions, Inc. for information on eucalyptus harvesting in Hawaiʻi. Thanks also to Christian Giardina and Zhiliang Zhu for making this project possible. Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S. Government.

>>>>

 Net primary production for each simulation cell was calculated as the mean IBIS-derived value for each combination of moisture zone and state type adjusted with a spatially explicit stationary growth multiplier to reflect local variation driven by microclimate [@sleeter_effects_2019]. We calculated this spatial growth multiplier as the NPP anomaly for each simulation cell relative to mean NPP values for each combination of moisture zone and state type based on empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] using Hawaiʻi-specific climate data [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014]. Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each combination of moisture zone and state type to the microclimate-adjusted NPP value for each simulation cell. All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each combination of moisture zone and state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018]. Climate change impacts on carbon flows were represented by temporal growth and decay multipliers applied to each simulation cell based on statistically downscaled CMIP5 climate projections for the Hawaiian Islands under each of two radiative forcing scenarios [RCP 4.5 and RCP 8.5; @timm_statistical_2015; @timm_future_2017]. The impact of future changes in rainfall and temperature on NPP were represented by annual growth multipliers calculated using empirical NPP models [@schuur_productivity_2003; @del_grosso_global_2008] and climate model projections of temperature and rainfall for each radiative forcing scenario. The effect of future warming on turnover rates of dead organic matter were represented by temporal decay multipliers calculated using Q10 functions and climate model temperature projections for each radiative forcing scenario. We applied a Q10 of 2.0 for wood and soil organic matter decay flows [@kurz_cbm-cfs3:_2009; @sleeter_effects_2019] and a Q10 of 2.17 for litter decay flows [@bothwell_leaf_2014]. Transition-triggered carbon flows resulting from disturbances associated with land use change, timber harvesting, and wildfire were based on values from @don_impact_2011, @selmants_baseline_2017, and @daniel_integrating_2018.

@@

### *4.2 Comparison to other studies*
There are few estimates of contemporary ecosystem carbon balance for the main Hawaiian Islands, and even fewer model projections of future ecosystem carbon balance in response to climate and land use change.  Our mean annual NBP estimate of 0.64 TgC y^-1^ for the period 2011-2020 agrees well with a recent State of Hawaiʻi Greenhouse Gas Inventory report, which estimated an annual net carbon sink of 0.66 TgC in 2015 from agriculture, forestry, and other land uses [@hawaii_inventory_2019]. In contrast, our NBP estimate for the past decade was ~ 88% higher than a previous statewide LUCAS model estimate covering the same time period [0.341 Tg C y^-1^; @selmants_baseline_2017]. This discrepancy was likely driven by modifications in how we calculated NPP, soil R~h~, and soil aquatic carbon loss compared to previous versions of the LUCAS model, as well our modelʻs finer spatial resolution [@selmants_baseline_2017; @daniel_integrating_2018]. Previous versions of a Hawaiʻi LUCAS model were run at 1-km spatial resolution and simulation cells within each unique combination of moisture zone and state type all had the same mean IBIS-derived NPP value applied to them at the beginning of the simulation period. In contrast, our NPP estimates at 250-m spatial resolution were adjusted on a cell-by-cell basis using Hawaiʻi-specific climate data as described in section 2.3. As a result, our statewide NPP estimates from 2011-2020 were 9.5% lower on average than previous LUCAS model estimates for Hawaiʻi during the same time period [@selmants_baseline_2017], likely because of the greater influence of more arid simulation cells. Soil carbon losses via R~h~, leaching, and overland flow in previous versions of the LUCAS model were calculated as the ratio of the IBIS-derived flux to the size of the originating carbon stock, in this case soil organic carbon to 1-m depth [@daniel_integrating_2018]. Here we calculated soil R~h~ and aquatic carbon losses as the ratio of the mean IBIS-derived flux to the microclimate-adjusted NPP value of each simulation cell, which is a more realistic driver than stock size [@jackson_ecology_2017]. Compared to previous Hawaiʻi LUCAS model estimates [@selmants_baseline_2017], soil R~h~ and aquatic carbon losses from 2011-2019 were reduced by an average of 15% and 21%, respectively, which widened the gap between between carbon gain (NPP) and carbon losses and accounted for the overall increase in NBP estimates for this time period.

@@

This study was funded by the U.S. Geological Survey Biological Carbon Sequestration Program. Thanks to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software, and to Nicholas Koch of Forest Solutions, Inc. for information on _Eucalyptus_ spp. harvesting in Hawaiʻi. Thanks also to Christian Giardina and Zhiliang Zhu for providing the initial impetus for this research. Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S. Government.
